### 2018-03-19 DiagrammeR/rmd_features.rmd
update

---
title: "Rmarkdown Features"
subtitle: "This is subtitle"
author: Yongfu Liao
date: March 18, 2018
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa-no-doi-no-issue.csl
link-citations: yes
output:
  html_document:
    css: ../style_sheets/style.css
    df_print: kable
    code_folding: show
    toc: true
    toc_depth: 4
    toc_float: true
---
<!--more-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg'
)
```

## yaml Setup {.tabset}

### Citation

Include the code in yaml header.

```{yaml}
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa-no-doi-no-issue.csl
link-citations: yes
```

### HTML Output Options

Include the code in yaml header.
```{yaml}
output:
  html_document:
    df_print: kable
    code_folding: hide
    fig_width: 7
    fig_height: 6
```

## Dynamic Variables

Include **R Code Chunks' Variables** in the body of text.

```{r}
alpha <- 1:3
text <- c("a", "b")
```

1. The numeric variable $\alpha =$ `r alpha[1]`.
    ```{yaml}
    # The numeric variable $\alpha =$ `r alpha[1]`.
    ```

2. The string variable $text =$ `r text[1]`.
    ```{yaml}
    # The string variable $text =$ `r text[1]`.
    ```

## $\LaTeX$ Rendering {.tabset}

### kable
```{r}
library(knitr)
library(dplyr)
iris_data <- head(iris) %>% select(Sepal.Length, Sepal.Width)

colnames(iris_data) <- c("$\\alpha$", "$\\frac{a}{b}$")
kable(iris_data, format = "markdown", align = "c")
```

### Plotting Expression
```{r fig.height=3}
library(latex2exp)
library(ggplot2)
iris_data2 <- iris[c(1:3, 50:53, 100 :103),] %>% select(Sepal.Length, Sepal.Width, Species)

iris_data2 %>% ggplot() +
    geom_point(mapping = aes(x=Sepal.Length, y=Sepal.Width, color=Species))+
    labs(x=TeX("Length($\\frac{kg}{cm^2}$)"),
         y=TeX("\\frac{a}{b}"))
```

## Citations 
To include **Citations**, [Read this](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html){target="_blank"}.

The text cited [@RN9].

@RN15 主張...

### Include `.bib` file in Reference {.tabset}

#### Specific References
Include specific references, which are **not cited** in the body of text.
```{pandoc}
---
nocite: |
  @item1, @item2
...
```

#### All References
Include **All references** in the body of text.
```{pandoc}
---
nocite: |
  @*
...
```

## Emphasis {.tabset}

### Full Example
```{html}
<div class="success">
| This is pragraph one. [**Red text**]{.red}, line 1. Text text text text.
| This is [line 2]{style="color:rgba(255, 0, 0, 1)"} of paragraph one.

Paragraph two starts here.
</div>
```

<div class="success">
| This is line 1, pragraph one. [**Red text**]{.red}. Text text text text.
| This is [line 2]{style="color:rgba(255, 0, 0, 1)"} of paragraph one.

Paragraph two starts here.
</div>

### Colored divs
There are 4 choices of classes, defined in my stylesheet[^css]:

1. success
2. warning
3. warning
4. error

```{html}
<div class="error">
some text in the colored region
</div>
```

<div class="error">
some text in the colored region
</div>

### New Line
```{markdown}
| Text of line 1
| Text of line 2

New paragraph
```

<div class="warning">
| Text of line 1
| Text of line 2

New paragraph
</div>

### Attach Attributes
By using `[`text`]`, pandoc turns the text in the square brackets into the HTML `<span>` tag. HTML `class`, `id`, attributes specific to `<span>` can then be attached by using the curly braces.

#### `class`
```{markdown}
[**Red Bold text**]{.red}
```
<div class="warning">
[**Red Bold text**]{.red}
</div>

Note[^red].

#### Other Attributes
```{markdown}
Add color through [span style]{style="color:rgba(255, 0, 0, 1)"}
```
<div class="warning">
Add color through [span style]{style="color:rgba(255, 0, 0, 1)"}
</div>

## Footnotes with Paragraphs
Check [Pandoc](http://pandoc.org/MANUAL.html#footnotes){target="_blank"} for details.

Look at a long footnote[^longnote]
```{markdown}
[^longnote]: Here's one with multiple blocks.

    Subsequent paragraphs are indented to show that they
belong to the previous footnote.

    In this way, multi-paragraph footnotes work like multi-paragraph list items.
```

## Shaped Images {.tabset}

### Rounded Edge
```{html}
<div class="rounded">
![](../icons/jekyll_dark.ico){width=auto height="30%"}
</div>
```
<div class="rounded">
![](../icons/jekyll_dark.ico){width=auto height="30%"}
</div>

### Circle Image
```{html}
<div class="circle">
![](../icons/jekyll_dark.ico){width=auto height="30%"}
</div>
```
<div class="circle">
![](../icons/jekyll_dark.ico){width=auto height="30%"}
</div>

<!-- footnotes -->
[^css]: Look at my [stylesheet](https://github.com/liao961120/local_depend/blob/master/style_sheets/style.css){target="_blank"} for syntax details.

[^longnote]: Here's one with multiple blocks.

    Subsequent paragraphs are indented to show that they
belong to the previous footnote.

    In this way, multi-paragraph footnotes work like multi-paragraph list items.

[^red]: `.red` is also a self-defined class in my stylesheet.
<!-- footnotes -->

<!-- syntax for Reference section --> 
## References