### 2018-03-24 1062_class/eco_model/HW/W2/HW2.rmd

---
title: "Ecological Modeling HW2"
author: Yongfu Liao
date: March 22, 2018
output: html_document
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```c
#include <stdio.h>

int main (void) {
    int age;

    printf("How old are you?\n");
    scanf("%d", &age);
    printf("You are %d years old.\n", age);

    return 0;
}
```
Download [source1.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/W2/source1.c){target="_blank"}

## Question 2

```c
#include <stdio.h>

int main (void) {
    int age;
    double height;

    printf("Please enter your age and height(cm).\n");
    scanf("%d", &age);
    scanf("%lf", &height);

    printf("Your age is %d and your height is %.1f cm.\n", age, height);

    return 0;
}
```
Download [source2.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/W2/source2.c){target="_blank"}