### 2018-03-29 1062_class/eco_model/HW/C4/HW3.rmd

---
title: "Ecological Modeling HW3"
author: Yongfu Liao
date: March 29, 2018
output: html_document
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

> Please input the bottom length of a triangle (integer). (output)
> 
> 3 ⏎ (input)
> 
> Please input the height of the triangle (integer). (output)
> 
> 4 ⏎ (input)
> 
> The area of the triangle is 6.00. (output)

```c
#include<stdio.h>

int main (void) {
    int bot_len, height;
    double area;

    printf("Please input the bottom length of a triangle (integer).\n");
    scanf("%d", &bot_len);

    printf("Please input the height of the triangle (integer).\n");
    scanf("%d", &height);

    area = (double) bot_len * (double) height / 2.0;
    printf("The area of the traingle is %.2lf.\n", area);

    return 0;
}

```
Download [Q1.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C4/Q1.c){target="_blank"}

## Question 2

> Please input your height (cm) and weight (kg). (output)
>
> 172.3 ⏎ (input)
>
> 65.0 ⏎ (input)
>
> Your BMI is XXXX. (output)
>
> The standard BMI is assumed to be 24.0. (output)
>
> The difference with a standard weight is YYY kg. (output)

```c
#include<stdio.h>

int main (void) {
    double height, weight, BMI;
    double height_meter, std_weight;

    printf("Please input your height (cm) and weight (kg).\n");
    scanf("%lf", &height);
    scanf("%lf", &weight);

    height_meter = height / 100.0;

    BMI = (weight) / (height_meter * height_meter);
    std_weight = 24.0 * (height_meter * height_meter);

    printf("Your BMI is %.2lf.\n", BMI);
    printf("The standard BMI is assumed to be 24.0.\n");
    printf("The difference with a standard weight is %.2lf kg. \n", weight - std_weight);

    return 0;
}
```
Download [Q2.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C4/Q2.c){target="_blank"}