### 2018-03-29 1062_class/eco_model/HW/C2/HW1.rmd
update

---
title: "Ecological Modeling HW1"
author: Yongfu Liao
date: March 15, 2018
output: html_document
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

> Correct the following source codes and arrange them in a readable way.

```c
#include <stdio.h>
int main (void) {Printf(“How are you? \n”); printf(“Not bad”\n):
return O;
```

```c
#include <stdio.h>

int main (void) {
	printf("How are you? \n"); 
	printf("Not bad\n");

	return 0;
}
```
Download [Q1.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C2/Q1.c){target="_blank"}

## Question 2

> Write a source code, which is intended to output the following tab
separated numbers.

```c
1.2 3.4 0.5
3.5 6.0 -10
```

```c
#include <stdio.h>

int main (void) {
	printf("%.1f\t%.1f\t%.1f\n", 1.2, 3.4, 0.5);
	printf("%.1f\t%.1f\t%d\n", 3.5, 6.0, -10);

	return 0;
}
```
Download [Q2.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C2/Q2.c){target="_blank"}