### 2018-04-02 1062_class/Multivariate_Analysis/HW2/Q3.Rmd
update

---
title: 'Q3: Air Pollution'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE,
                      dev='svg',
                      fig.dim=c(3.6, 3.2))
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
theme <- theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
```

```{r}
air<- read_table2("data/T1-5.dat", col_names = FALSE)
colnames(air) <- c("wind", "radiation", "CO", "NO", "NO2", "O3", "HC")
```

PCA with Covariance Matrix vs. Correlation Matrix
================================
```{r}
PCA_cov <- prcomp(air, scale=F)
PCA_cor <- prcomp(air, scale =T)
```

```{r fig.dim=c(6, 3.3)}
library(factoextra)
library(cowplot)

p_cov <- fviz_eig(PCA_cov, choice = "eigenvalue", main="PCA with S", geom = "line")
p_cor <- fviz_eig(PCA_cor, choice = "eigenvalue", main="PCA with R", geom = "line")

plot_grid(p_cov, p_cor, nrow = 1)
```

### Principle Components(eigenvector) of Covariance Matrix
```{r}
kable(PCA_cov[["rotation"]], format = "markdown", align = "c", digits = 2)
```

### Principle Components(eigenvector) of Correlation Matrix
```{r}
kable(PCA_cor[["rotation"]], format = "markdown", align = "c", digits=2)
```