### 2018-04-02 1062_class/Multivariate_Analysis/HW2/Q6.Rmd

---
title: 'Q6: questionare'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE)
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r data_import_rm.na}
var_content <- read_delim("data/var_masst.txt", 
    ";", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)
colnames(var_content) <- NULL

response <- read.table("C:/Users/user/local_depend/1062_class/Multivariate_Analysis/HW2/data/masst.txt", quote="\"", comment.char="", na.strings=".", stringsAsFactors=FALSE) %>%
    drop_na()
```

```{r}
PCA <- prcomp(response, scale = TRUE)

std_pc <- PCA[["sdev"]] 
# std(sqrt(eigenvalue) of corr matrix) of PCs
```

(a)
===============================

### Loading of the First 5 Pricinple Components
```{r}
# First 5 PCs
pc5 <- as.data.frame(PCA[["rotation"]][,1:5])

styled_pc5 <- round(pc5,3) %>%
    mutate(
        PC1 = cell_spec(PC1, "html", color = ifelse(PC1 < 0, "red", "blue")),
        PC2 = cell_spec(PC2, "html", color = ifelse(PC2 < 0, "red", "blue")),
        PC3 = cell_spec(PC3, "html", color = ifelse(PC3 < 0, "red", "blue")),
        PC4 = cell_spec(PC4, "html", color = ifelse(PC4 < 0, "red", "blue")),
        PC5 = cell_spec(PC5, "html", color = ifelse(PC5 < 0, "red", "blue"))
                        )

loadings <- kable(styled_pc5, "html", escape = F, align="c", row.names = T) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
loadings
```

### Percentage of Variance Accounted
```{r}
per_5pc <- sum(std_pc[1:5]^2)/sum(std_pc^2)
```
The percentage accounted is `r round(per_5pc*100,2)`%.

(b)
==============================

```{r fig.show='hold'}
loadings
var_content %>% 
kable(format = "html", align="l") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```