### 2018-04-03 1062_class/Multivariate_Analysis/HW2/all.Rmd

---
title: "MultiAnaly HW2"
output: html_document
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE,
                      dev='svg',
                      fig.dim=c(3.6, 3.2))
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(mat2tex)
theme <- theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
source("multivariate_fc.R")
```

Q1: 8.1
====================================

```{r}
covmt <- matrix(c(5,2,2,2), nrow = 2)
eig <- eigen(covmt)
```

```{r}
eigval <- eig$values
eigvec <- eig$vectors
percent <- eigval[1]/sum(eigval)
```

- First Principle Component: 
    $\lambda_1 = `r eigval[1]`$, $Y_1 = `r round(eigvec[1,1], 3)`X_1 `r round(eigvec[2,1], 3)`X_2$

- Second Principle Component:
    $\lambda_2 = `r eigval[2]`$, $Y_2 = `r round(eigvec[1,2], 3)`X_1 `r round(eigvec[2,2], 3)`X_2$

- Proprotion of variance explained by the first PC: 
    $\frac{\lambda_1}{\Sigma \lambda_i} = `r round(percent*100, 2)`$%

Q2: 8.2
=======================================

```{r}
cormt <- diag(1/sqrt(diag(covmt)))%*%covmt%*%diag(1/sqrt(diag(covmt)))
```

$\rho = diag(\Sigma)^{\frac{-1}{2}} ~ \Sigma ~ diag(\Sigma)^{\frac{-1}{2}} =
```{r results='asis'} 
xm(cormt, round=T) %_% "$"
```

### (a)
```{r}
eig <- eigen(cormt)
eigval <- eig$values
eigvec <- eig$vectors
percent <- eigval[1]/sum(eigval)
```

- First Principle Component: 
    $\lambda_1 = `r round(eigval[1],2)`$, $Y_1 = `r round(eigvec[1,1], 3)`Z_1 `r round(eigvec[2,1], 3)`Z_2$

- Second Principle Component:
    $\lambda_2 = `r round(eigval[2],2)`$, $Y_2 = `r round(eigvec[1,2], 3)`Z_1 `r round(eigvec[2,2], 3)`Z_2$

- Proprotion of variance explained by the first PC: 
    $\frac{\lambda_1}{\Sigma \lambda_i} = `r round(percent*100, 2)`$%

### (b)

Principle components from 8.1 & 8.2 are different since they came from two different matrix (covariance matrix vs. correlation matrix). Mathmetically, the difference resulted from two different matrices have different eigenvalues and eigenvectors, hence different principle component. This difference reflects the fact that scaling has an effect on the principle component.

### (c)

```{r}
e_11 <- round(eigvec[1,1],3)
sqrt_eigen_1 <- round(sqrt(eigval[1]), 3)
v11 <- eigvec[1,1]*sqrt(eigval[1])

e_12 <- round(eigvec[2,1],2)
v12 <- eigvec[2,1]*sqrt(eigval[1])

e_21 <- round(eigvec[1,2],2)
sqrt_eigen_2 <- round(sqrt(eigval[2]), 2)
v21 <- eigvec[1,2]*sqrt(eigval[2])
```

$\rho_{Y_i,~Z_k} = \frac{e_{ik} \sqrt{\lambda_i}} {\sqrt{\sigma_{kk}}}$

$\rho_{Y_1,~Z_1} = \frac{(`r e_11`)(`r sqrt_eigen_1`) }{\sqrt{1}} = `r round(v11,3)`$

$\rho_{Y_1,~Z_2} = `r round(v12,3)`$

$\rho_{Y_2,~Z_1} = `r round(v21,3)`$

Q3: PCA with Covariance Matrix vs. Correlation Matrix
================================
```{r}
air<- read_table2("data/T1-5.dat", col_names = FALSE)
colnames(air) <- c("wind", "radiation", "CO", "NO", "NO2", "O3", "HC")
```

```{r}
PCA_cov <- prcomp(air, scale=F)
PCA_cor <- prcomp(air, scale =T)
```

```{r fig.dim=c(6, 3.3)}
library(factoextra)
library(cowplot)

p_cov <- fviz_eig(PCA_cov, choice = "eigenvalue", main="PCA with S", geom = "line")
p_cor <- fviz_eig(PCA_cor, choice = "eigenvalue", main="PCA with R", geom = "line")

plot_grid(p_cov, p_cor, nrow = 1)
```

### Principle Components(eigenvector): Covariance Matrix
```{r}
kable(PCA_cov[["rotation"]], format = "markdown", align = "c", digits = 2)
```

### Principle Components(eigenvector): Correlation Matrix
```{r}
kable(PCA_cor[["rotation"]], format = "markdown", align = "c", digits=2)
```

Q4
==================================
```{r data_import}
sweet <- read_table2("data/T5-1.txt", col_names = FALSE)
colnames(sweet) <- c("sweet_rate", "sodium", "potassium")
```

```{r fig.dim=c(7, 4)}
source("multivariate_fc.R")
PCA <- prcomp(sweet, scale=F)
PCA_cor <- prcomp(sweet, scale=T)

p_cov <- scree_plot(PCA, "variance", addlabels=TRUE, main="Covariance mat", sub_axis_name="", axis_limits = c(0,100))
p_cor <- scree_plot(PCA_cor, "variance", addlabels=TRUE, main="Correlation mat",ylab = "", axis_limits = c(0,100))
cowplot::plot_grid(p_cov, p_cor, nrow = 1)
```

```{r}
library(mat2tex)
kable(PCA[["rotation"]], format="markdown", align="c", digits = 2)
```

```{r}
quantiles <- qnorm(p=seq(0.25,9.75, 0.5)/10) 
PC1 <- as.data.frame(PCA[["x"]][,1])
    PC1 <- cbind(rownames(PC1), PC1)
    colnames(PC1) <- c("ID", "PC1")
    PC1 <- PC1 %>% 
        arrange(PC1) %>% mutate(quantile=quantiles)

PC2 <- as.data.frame(PCA[["x"]][,2])
    PC2 <- cbind(rownames(PC2), PC2)
    colnames(PC2) <- c("ID", "PC2")
    PC2 <- PC2 %>% 
        arrange(PC2) %>% mutate(quantile=quantiles)

PC3 <- as.data.frame(PCA[["x"]][,3])
    PC3 <- cbind(rownames(PC3), PC3)
    colnames(PC3) <- c("ID", "PC3")
    PC3 <- PC3 %>% 
        arrange(PC3) %>% mutate(quantile=quantiles)
```

```{r fig.dim=c(7.5, 3.3)}
library(ggplot2)
library(scales)
library(cowplot)
library(latex2exp)

p1 <- ggplot(PC1, aes(x=quantiles, y=PC1))+
        geom_point(color=hue_pal()(3)[1])+
        geom_text(aes(label=ID),
                  hjust=-0.1, vjust=0, size = 2)+
        scale_x_continuous(breaks = seq(-2,2, 0.5))+
        scale_y_continuous(breaks = seq(-35,35, 5))+
        labs(x=TeX("$q_{(j)}$"),
             y="PC1",
             title="PC1 Q-Q Plot")+ theme

p2 <- ggplot(PC2, aes(x=quantiles, y=PC2))+
        geom_point(color=hue_pal()(3)[2])+
        geom_text(aes(label=ID),
                  hjust=-0.1, vjust=0, size = 2)+
        scale_x_continuous(breaks = seq(-2,2, 0.5))+
        scale_y_continuous(breaks = seq(-4.5,3.5, 0.5))+
        labs(x=TeX("$q_{(j)}$"),
             y="PC2",
             title="PC2 Q-Q Plot")+ theme

p3 <- ggplot(PC3, aes(x=quantiles, y=PC3))+
        geom_point(color=hue_pal()(3)[3])+
        geom_text(aes(label=ID),
                  hjust=-0.1, vjust=0, size = 2)+
        scale_x_continuous(breaks = seq(-2,2, 0.5))+
        scale_y_continuous(breaks = seq(-2,2, 0.5))+
        labs(x=TeX("$q_{(j)}$"),
             y="PC3",
             title="PC3 Q-Q Plot")+ theme

plot_grid(p1, p2, p3, nrow = 1)
```

Q5
===================================
```{r data_import2}
fp <- read_table2("data/FOODP.txt", col_names = FALSE, 
    skip = 5)[1:24, -1]
colnames(fp) <- c("city", paste("food",1:5,sep=""))
```

### (a) Scree Plot, PCs, Interpretation

```{r scree_plot}
library(factoextra) # For Visualizing MvA
PCA <- prcomp(fp[,-1], scale=T)
# fviz_eig(PCA)

eigva <- as_data_frame(PCA[["sdev"]]^2)
scree <- cbind(1:5 , eigva)
scree <- cbind(scree, "a")
colnames(scree) <- c("PC", "var", "a")

ggplot(scree, aes(x = PC, y = var, group = a)) +
    geom_point() + geom_line() +
    scale_y_continuous(
        sec.axis = sec_axis(~ . / sum(eigva) * 100,
                            name = "Variance Accounted (%)")
                   )+
    labs(x="Principle Components",
         y=latex2exp::TeX("$\\hat{\\lambda_i}$"),
         title="Scree Plot")    
```

By either the "elbow" of the scree plot or the unity criterion, **one principle component** is sufficient, which explains **`r round(eigva[1,1]/sum(eigva)*100, 2)`%** of the total variance.

```{r PC_loadings}
pc <- PCA[["rotation"]]
kable(round(pc,3),"markdown", align = "c")
```

### (b) The First Two principal component scores

```{r}
# PCA[["x"]] returns the "rotated" score(pc score) of the original data
pc_score12 <- as.data.frame(PCA[["x"]][,1:2])
pc_score12 <- cbind(fp$city, pc_score12) %>% rename(city=`fp$city`)
kable(pc_score12, "markdown", align="c",digits = 2)
```

### (c) Scatter Plot: PC1 vs. PC2

```{r fig.dim=c(7, 4)}
ggplot(pc_score12, aes(x=PC1, y=PC2))+
    geom_point()+
    geom_text(aes(label=city),
              hjust=-0.1, vjust=0, size = 3)+
    scale_x_continuous(limits = c(-4.6, 2.9), breaks = seq(-4.5,3, 0.5))+
    scale_y_continuous(limits = c(-1.7, 1.5), breaks = seq(-1.75,1.5, 0.25))+
    labs(title="Scatter Plot of PC1 & PC2",
         x="Principle Component 1",
         y="Principle Component 2")
```

### (d) Food Price in Different Cities
```{r}
## Max PC score: cheapest food price
cheap <- pc_score12$city[pc_score12$PC1 == max(pc_score12$PC1)]

## min PC score
expense <- pc_score12$city[pc_score12$PC1 == min(pc_score12$PC1)]
```

Using principle component 1 as indicator, since it has the same sign on all "weighted" food price, and it's the most important principle component in explaining the variability of the data, the x-axis of the scatter above showed that **`r cheap`** has the cheapest food price, and **`r expense`** has the most expensive food price.

Q6
================================
```{r data_import_rm.na}
var_content <- read_delim("data/var_masst.txt", 
    ";", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)
colnames(var_content) <- NULL

response <- read.table("C:/Users/user/local_depend/1062_class/Multivariate_Analysis/HW2/data/masst.txt", quote="\"", comment.char="", na.strings=".", stringsAsFactors=FALSE) %>%
    drop_na()
```

```{r}
PCA <- prcomp(response, scale = TRUE)

std_pc <- PCA[["sdev"]] 
# std(sqrt(eigenvalue) of corr matrix) of PCs
```

### (a)

#### Loading of the First 5 Pricinple Components
```{r}
# First 5 PCs
pc5 <- as.data.frame(PCA[["rotation"]][,1:5])

styled_pc5 <- round(pc5,3) %>%
    mutate(
        PC1 = cell_spec(PC1, "html", color = ifelse(PC1 < 0, "red", "blue")),
        PC2 = cell_spec(PC2, "html", color = ifelse(PC2 < 0, "red", "blue")),
        PC3 = cell_spec(PC3, "html", color = ifelse(PC3 < 0, "red", "blue")),
        PC4 = cell_spec(PC4, "html", color = ifelse(PC4 < 0, "red", "blue")),
        PC5 = cell_spec(PC5, "html", color = ifelse(PC5 < 0, "red", "blue"))
                        )

loadings <- kable(styled_pc5, "html", escape = F, align="c", row.names = T) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
loadings
```

#### Percentage of Variance Accounted
```{r}
per_5pc <- sum(std_pc[1:5]^2)/sum(std_pc^2)
```
The percentage accounted is `r round(per_5pc*100,2)`%.

### (b)

```{r fig.show='hold'}
var_content %>% 
kable(format = "html", align="l") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)