### 2018-04-12 Instructions/Rmarkdown_features/rmd_features.rmd

---
title: "Rmarkdown Features"
subtitle: "This is subtitle"
author: Yongfu Liao
date: "`r Sys.Date()`"
bibliography: /Users/user/local_depend/citation/neanderthals.bib
csl: /Users/user/local_depend/style_sheets/apa-no-doi-no-issue.csl
link-citations: yes
output:
  html_document:
    code_folding: show
    css: /Users/user/local_depend/style_sheets/style.css
    df_print: kable
    toc: true
    toc_depth: 3
    toc_float:
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg',
	fig.dim = c(5.5, 3.5)  # (width, height)
)
```

yaml Setup {.tabset}
-----------------------------

yaml metadata determines the output file format, looking, and any other additional customizing features of Rmarkdown files.

### yaml of this Doc

This is the complete yaml header of this document.

```yaml
---
title: "Rmarkdown Features"
subtitle: "This is subtitle"
author: Yongfu Liao
date: March 18, 2018
bibliography: /Users/user/local_depend/citation/neanderthals.bib
csl: /Users/user/local_depend/style_sheets/apa-no-doi-no-issue.csl
link-citations: yes
output:
  html_document:
    code_folding: show
    css: /Users/user/local_depend/style_sheets/style.css
    df_print: kable
    toc: true
    toc_depth: 3
    toc_float:
        collapsed: false
---
```

### HTML Output

```{yaml}
output:
  html_document:
    css: ../style_sheets/style.css
    df_print: kable
    code_folding: hide
    fig_width: 7
    fig_height: 6
    toc: true
    toc_depth: 4
    toc_float: true
```

### PDF Output

You may need to attatch an additional `header.tex` and set `latex_engine` to `xelatex` for PDF documents containing multibite characters (e.g. Chinese). Read more about  [header.txt](#pdf-header).
```yaml
---
output:
  pdf_document:
    latex_engine: xelatex    
    includes:
        in_header: header.tex
documentclass: article
urlcolor: "blue"
fontsize: 12pt
---
```

### PDF: header.tex{#pdf-header}

The code below is for generating PDF documents written in (Traditional) Chinese.

```tex
%加這個就可以設定字體
\usepackage{fontspec}
%使用xeCJK，其他的還有CJK或是xCJK
\usepackage{xeCJK}

%設定英文字型，不設的話就會使用預設的字型
\setmainfont{Calibri}

%設定中英文的字型
%字型的設定可以使用系統內的字型，而不用像以前一樣另外安裝
\setCJKmainfont{蘋方-繁}

%中文自動換行
\XeTeXlinebreaklocale "zh"

%文字的彈性間距
\XeTeXlinebreakskip = 0pt plus 1pt

%設定段落之間的距離
\setlength{\parskip}{0.15cm}

%設定行距
\linespread{1.2}\selectfont
```

### Bibliography

```{yaml}
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa-no-doi-no-issue.csl # APA style sheet
link-citations: yes
```

For citation syntax, read the section [Bibliography & Citations](#cite).

Dynamic Variables
-----------------------------

Include **R Code Chunks' Variables** in the body of text.

```{r}
alpha <- 1:3
text <- c("a", "b")
```

1. The numeric variable $\alpha =$ `r alpha[1]`.
    ```{yaml}
    # The numeric variable $\alpha =$ `r alpha[1]`.
    ```

2. The string variable $text =$ `r text[1]`.
    ```{yaml}
    # The string variable $text =$ `r text[1]`.
    ```

$\LaTeX$ Math Rendering {.tabset}
-----------------------------
### Passing Matrix Object

#### Setup

Add `results='asis'` to r code chunk option.
```{r results='asis'}
library(mat2tex)
mt <- c(1,2)
"$A =" %_% xm(mt,0) %_%"_{2 \\times 1}$"
```

#### Include Matrix in Text

![](mat2tex.png){width=100% height=auto}

renders as:

<div class="success">
This is the variable `mt`,
```{r results='asis', echo=FALSE}
library(mat2tex)
mt <- matrix(1:4, 2,2)
"$A =" %_% xm(mt,0) %_%"^{-1}$"
```
, which is a matrix.
</div>

Check [here](http://markheckmann.github.io/mat2tex/){target=_blank""} for more information.

### kable
```{r}
library(knitr)
library(dplyr)
iris_data <- head(iris) %>%
    select(Sepal.Length, Sepal.Width)

colnames(iris_data) <- c("$\\alpha$", "$\\frac{a}{b}$")
kable(iris_data, format = "markdown", align = "c")
```

### Plotting Expression
```{r fig.height=3}
library(latex2exp)
library(ggplot2)
iris_data2 <- iris[c(1:3, 50:53, 100:103), ] %>%
    select(Sepal.Length, Sepal.Width, Species)

iris_data2 %>% ggplot() +
    geom_point(aes(x = Sepal.Length, 
                   y = Sepal.Width, 
                   color = Species)) +
    labs(x = TeX("Length($\\frac{kg}{cm^2}$)"),
         y = TeX("\\frac{a}{b}"))
```

Bibliographies & Citations{#cite} 
-----------------------------

You can read the full documentation of [Bibliographies & Citations](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html){target="_blank"} at Rstudio.

### Citation syntax

#### Classical Citation: **Author**, **year**, and **brackets**
- `Some text [@Bib.ID].`
    - Some text [@RN9].

#### In-text Citation: **Author (year)**
- `@Bib.ID Some text`
    - @RN15 Some text

- `@Bib.ID [p. 20] Some text.`
    - @RN9 [p. 20] Some text.

#### Year only: Suppress Author
- `Some text [-@Bib.ID].`
    - Some text [-@RN9]

#### Two Citations
- `Some text [@Bib.ID1; @Bib.ID2].`
    -  Some text [@RN9; @RN15].

5. Prefix & Suffix
    - `Some text [see @Bib.ID1 pp.45; also, @Bib.ID2 ch. 2].`
        -  Some text [see @RN9 pp. 45; also, @RN15 ch. 2].

### Include `.bib` file in Reference {.tabset}

Rmarkdown requires specific formats of bibliographies. One of the most common one is `.bib`. A lot of people (including me) use Endnote to manage bibliographies and have no time getting used to other open source reference managers. Read  [this](https://libguides.usask.ca/c.php?g=218034&p=1458583) to see how to generate `.bib` from EndNote.

Below are some syntax in Rmarkdown to control the display of reference in the document.

#### Reference heading

The body of cited reference in the `.bib` file will be shown under the header named Reference (it doesn't matter which **level** of heading you use).

```markdown
## Reference
```

#### Specific Refs
Include specific references that were **not cited** in the body of text.
```{pandoc}
---
nocite: |
  @item1, @item2
...
```

#### All Refs
Include **All references** in the body of text.
```{pandoc}
---
nocite: |
  @*
...
```

Emphasis {.tabset}
-----------------------------

### An Example

````markdown
`r ''````{block2, type="success"}
| This is line 1, pragraph one. [**Red text**]{.red}. Text text text text.
| This is [line 2]{style="color:rgba(255, 0, 0, 1)"} of paragraph one.

Paragraph two starts here.
```
````

```{block2, type="success"}
| This is line 1, pragraph one. [**Red text**]{.red}. Text text text text.
| This is [line 2]{style="color:rgba(255, 0, 0, 1)"} of paragraph one.

Paragraph two starts here.
```

### Colored divs

#### Bootstrap Alert

You can use bootstrap's default alert divs in Rmarkdown, defined with four classes:

1. alert alert-success
2. alert alert-warning
3. alert alert-info
4. alert alert-danger

````markdown
`r ''````{block, type="alert alert-success"}
Bootstrap-defined colored region
```
````

```{block, type="alert alert-success"}
Bootstrap-defined colored region
```

#### Customized divs

Alternative, you can define your own styled divs, as in my stylesheet[^css]:

There are 4 choices of classes,

1. success
2. warning
3. info
4. error

````markdown
`r ''````{block, type="info"}
some text in the colored region
```
````

```{block, type="info"}
some text in the colored region
```

### Bootstrap Callouts

Althougt it's "Bootstrap" callouts, it's not contained in default bootstrap theme. I defined it in my stylesheet.

There are 4 choices of classes,

1. bs-callout bs-callout-success
2. bs-callout bs-callout-warning
3. bs-callout bs-callout-info
4. bs-callout bs-callout-error

````markdown
`r ''````{block2, type="bs-callout bs-callout-warning"}
#### Heading in Callout
**Calleout Text** texttexttexttexttexttext
```
````

```{block2, type="bs-callout bs-callout-warning"}
#### Heading in Callout
**Calleout Text** texttexttexttexttexttext
```

### New Line
```{markdown}
| Text of line 1
| Text of line 2

New paragraph
```

```{block2, type="alert alert-success"}
| Text of line 1
| Text of line 2

New paragraph
```

### Attach Attributes
By using `[`text`]`, pandoc turns the text in the square brackets into the HTML `<span>` tag. HTML `class`, `id`, attributes specific to `<span>` can then be attached by using the curly braces.

#### `class`

Rmarkdown syntax allows attaching attributes (such as class, id, etc.) to certain elements by wrapping attributes into `{ ... }`.

Elements allowed attributes to be attach (by markdown syntax directly):

1. Section: Attributes are added next to **any level of heading**.
2. Link: `[Text](LINK){attributes}`
3. `span`: Text wrapped in `[`, `]`. e.g `[Text with class "red"]{.red}`.

```{markdown}
[**Red Bold text**]{.red}
```

```{block2, type="alert alert-success"}
[**Red Bold text**]{.red}
```

Note[^red].

#### Other Attributes
```{markdown}
Add color through [span style]{style="color:rgba(255, 0, 0, 1)"}
```

```{block2, type="alert alert-success"}
Add color through [span style]{style="color:rgba(255, 0, 0, 1)"}
```

Footnotes with Paragraphs
-----------------------------
Check [Pandoc](http://pandoc.org/MANUAL.html#footnotes){target="_blank"} for details.

```{markdown}
[^longnote]: Here's one with multiple blocks.

    Subsequent paragraphs are indented to show that they belong to the previous footnote. (JUST IGNORE ME, I AM Exteeeeeendiiiiiiing the text to show the effect of paragraph).

    In this way, multi-paragraph footnotes work like multi-paragraph list items.
```
- Read the output by clicking the superscript[^longnote].

Shaped Images {.tabset}
-----------------------------

### Rounded Edge
```{markdown}
![](../../local_depend/icons/libreoffice.ico){width=auto height="30%" .rounded}
```
![](/Users/user/local_depend/icons/libreoffice.ico){width=auto height="30%" .rounded}

#### Boostrap
```{markdown}
![](../../local_depend/icons/libreoffice.ico){width=auto height="30%" .img-rounded}
```
![Bootstrap Rounded](/Users/user/local_depend/icons/libreoffice.ico){width=auto height="30%" .img-rounded}

### Circle Image
```{markdown}
![](../../local_depend/icons/libreoffice.ico){width=auto height="30%" .circle}
```
![This is figure Caption](/Users/user/local_depend/icons/libreoffice.ico){width=auto height="30%" .circle}

#### Boostrap
```{markdown}
![](../../local_depend/icons/libreoffice.ico){width=auto height="30%" .img-circle}
```
![Bootstrap Circle](/Users/user/local_depend/icons/libreoffice.ico){width=auto height="30%" .img-circle}

### Thumbnail
```{markdown}
![](../../local_depend/icons/libreoffice.ico){width=auto height="30%" .img-thumbnail}
```
![Bootstrap Thumbnail](/Users/user/local_depend/icons/libreoffice.ico){width=auto height="30%" .img-thumbnail}

<!-- footnotes ---------------------------- -->
[^css]: Look at my [stylesheet](https://github.com/liao961120/local_depend/blob/master/style_sheets/style.css){target="_blank"} for syntax details.

[^longnote]: Here's one with multiple blocks.

    Subsequent paragraphs are indented to show that they belong to the previous footnote. (JUST IGNORE ME, I AM Exteeeeeendiiiiiiing the text to show the effect of paragraph).

    In this way, multi-paragraph footnotes work like multi-paragraph list items.

[^red]: `.red` is also a self-defined class in my stylesheet.
<!-- footnotes -------------------------- -->

<!-- syntax for Reference section -->

Figures
-----------------------------

### Captions
Include `fig.cap='An Caption of the Plot'` in the code chunk option gives:
```{r fig.cap='An Caption of the Plot'}
library(ggplot2)
ggplot(iris)+
    geom_point(aes(x=Sepal.Length, y=Sepal.Width, color=Species))
```

References
-----------------------------