### 2018-04-20 1062_class/eco_model/HW/C5/HW4.rmd
update

---
title: "Ecological Modeling HW4"
author: Yongfu Liao
date: April 12, 2018
output: html_document
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

> Please input an integer number. (output)
>
> 3 (or 4) ⏎ (input)
>
> 3 (or 4) is an odd (or even) number . (output)

```c
#include<stdio.h>

int main (void) {
    int input;
    int i;

    printf("Please input an integer.\n");
    scanf("%d", &input);

    if (input % 2 == 0) {
        printf("%d is an even number\n", input);
    } else {
        printf("%d is an odd number\n", input);
    }

    return 0;
}
```
Download [even_odd.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C5/even_odd.c){target="_blank"}

## Question 2

> Please input your height (cm) and weight (kg). (output)
>
> 172.3 ⏎ (input)
>
> 65.0 ⏎ (input)
>
> Your BMI is 21.890. (output)
>
> The standard BMI is assumed to be 24.0. (output)
>
> I’m afraid you are overweight (or underweight). (output)

```c
#include<stdio.h>

int main (void) {
    double height, weight, BMI;
    double height_meter;

    printf("Please input your height (cm) and weight (kg).\n");
    scanf("%lf", &height);
    scanf("%lf", &weight);

    height_meter = height / 100.0;

    BMI = (weight) / (height_meter * height_meter);

    printf("Your BMI is %.2lf.\n", BMI);
    printf("The standard BMI is assumed to be 24.0.\n");

    if (BMI > 24.0) {
        printf("I’m afraid you are overweight.\n");
    } else {
        printf("I’m afraid you are underweight.\n");
    }

    return 0;
}
```
Download [BMI.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C5/BMI.c){target="_blank"}

## Question 3

> Please input an integer number (> 1). (output)
> 
> 7 ⏎ (input)
> 
> 7 is a prime number. (output)
> 
> Please input an integer number (> 1). (output)
> 
> 15 ⏎ (input)
>
> 15 is not a prime number. (output)

```c
#include<stdio.h>

int main (void) {
    int input;
    int i;

    printf("Please input an integer number (> 1).\n"); 
    scanf("%d", &input);

    for (i=2; i<=input; i++){
        if (input % i == 0 && input != i) {
            printf("%d is not a prime number.\n", input);
            break;
        }//end if
    }//end for

    if (i == input+1){
        printf("%d is a prime number.\n", input);
    }//end if
    return 0;
}

```
Download [prime.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C5/prime.c){target="_blank"}