### 2018-04-22 1062_class/Multivariate_Analysis/HW3_factor_analysis/HW3.Rmd

---
title: "多變量分析 HW3"
author: "劉昱維, 吳冠瑋, 廖永賦, 謝靖惟, 黃奎鈞"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: header.tex
css: /Users/user/local_depend/style_sheets/style.css
always_allow_html: yes
---

```{r setup, include=FALSE}
if (knitr::is_html_output()) {
    knitr::opts_chunk$set(echo = FALSE,
                          message=FALSE,
                          warning=FALSE,
                          comment = "",
                          dev='svg',
                          fig.dim=c(3.6, 3.2))
} else {
    knitr::opts_chunk$set(echo = FALSE,
                          message=FALSE,
                          warning=FALSE,
                          comment = "",
                          out.width = '70%',
                          out.height = '5.5cm')
}
options(digits = 4)
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(mat2tex)
theme <- theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
source("multivariate_fc.R")
```

Q1: 9.12
====================================

```{r Covariance_matrices}
S <- 10^(-3)*matrix(c(11.072, 8.019, 8.160,
                      8.019, 6.417, 6.005,
                      8.160, 6.005, 6.773), 
                    nrow = 3, ncol = 3)

Sn <- (23/24)*S # n = 24, Convert to Sn for M.L.E.
```

```{r results='asis', echo=FALSE}
"$\\mathbf{S_n} = \\frac{23}{24} \\mathbf{S} =" %_% xm(Sn ,5) %_% "$"
```

## (a) Specific Variances

```{r}
L <- c(.1022, .0752, .0765)
LL <- matrix(L, ncol = 1) %*% matrix(L, nrow = 1)
Psi <- diag(Sn - matrix(L, ncol = 1) %*% matrix(L, nrow = 1))
```

$\mathbf{S_n} \approx \hat{\mathbf{L}} \hat{\mathbf{L}}^T + \hat{\mathbf{\Psi}}$, where $diag(\mathbf{S_n}) = diag(\hat{\mathbf{L}} \hat{\mathbf{L}}^T) + diag(\hat{\mathbf{\Psi}})$

Hence, $\hat{\mathbf{\Psi}} = diag(\hat{\mathbf{\Psi}}) = diag(\mathbf{S_n} - \hat{\mathbf{L}} \hat{\mathbf{L}}^T)$

```{r results='asis', echo=FALSE}
"$\\hat{\\mathbf{\\Psi}} = " %_% xm(diag(Psi), 6) %_% "$"
```

## (b) Communalities

$\sigma_{ii} = \ell_{i1}^2 + \ell_{i2}^2 + ~...~ + \ell_{im}^2 + \psi_i$

$h_i^2 = \ell_{i1}^2 + \ell_{i2}^2 + ~...~ + \ell_{im}^2 = `r sum(L^2)`$

## (c) Proportion of variance explained by the factor

```{r}
p <-  sum(L^2) / (diag(Sn) %>% sum())
```

$\frac{s_{11} + s_{22} + ~...~ + s_{pp}}{h_i^2} = `r p`$

## (d) Residual Matrix

```{r}
Residual <- Sn - LL - Psi 
```

$\mathbf{S_n} - \mathbf{\hat{L}}\mathbf{\hat{L}}^T - \mathbf{\hat{\Psi}} = 
```{r results='asis'} 
xm(Residual, 6) %_% "$"
```

Q2: 9.32
====================================
```{r}
library(readr)
library(psych)
bulls <- read_table2("T1-10.dat", col_names = FALSE)
colnames(bulls) <- c("Breed", "SalePr", "YrHgt", 
                     "FtFrBody", "PrctFFB", "Frame",
                     "BkFat", "SaleHt", "SaleWt")
bulls <- bulls[,-c(1,9)]
```

### (a) S PC
```{r}
fa_3 <- fa(cov(bulls), nfactors=3, rotate="varimax", covar = T, fm="pa", scores="regression")
#no principle component method
```

### (b) S MLE

```{r}
fa_3 <- fa(cov(bulls), nfactors=3, rotate="varimax", covar = T, fm="ml", scores="regression")
```

### (c) R PC

```{r}
fa_3 <- fa(cov(bulls), nfactors=3, rotate="varimax", covar = T, fm="pa", scores="regression")
#no principle component method
```

### (d) R MLE
```{r}
fa_1 <- factanal(bulls, factors = 1, scores = "regression", rotation = "varimax")
fa_2 <- factanal(bulls, factors = 2, scores = "regression", rotation = "varimax")
fa_3 <- factanal(bulls, factors = 3, scores = "regression", rotation = "varimax")
```

### (e) Compare

### (f) scatter plots of factor2 vs factor1 in (a) and (c)

Q3
====================================

```{r}
library(readr)
drinks <- read_table2("drinks.DAT", col_names = FALSE, skip = 2)
colnames(drinks) <- c("ID", "BRAND",
                      paste("X",1:10, sep=""))
```

```{r}
library(psych)
items <- drinks[,-(1:2)]
fa(cov(items), nfactors=3, rotate="varimax", covar = F, fm="pa", scores="regression")
```

#no principle component method