### 2018-04-22 1062_class/eco_model/HW/C6/HW5.rmd

---
title: "Ecological Modeling HW5"
author: Yongfu Liao
date: "`r Sys.Date()`"
output: 
    html_document:
        code_folding: hide
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
	)
```

## Question

Bifurcation:

$$N_{t+1} = N_t ~ e^{r (1-N_t)}, ~~1 \leq r  \leq 4.5 $$

### C Code
```c
#include <stdio.h>
#include <math.h>

int main (void) {

    //i_max determines the length of interval of r
    int i, i_max=175;

    //j_max determines the number of iterations under each r
    int j, j_max=1000;

    double r = 1; // population growth potential
    double N_int=0.1, N = N_int; // population size

    printf("N\tr\n");
    for (i=0; i<=i_max; i++){
        for (j=0; j<=j_max; j++){ //under specific r
            N = N * exp(r*(1 - N)); 
            if (j > j_max/1.5) { // print values only at final phase
                printf("%lf\t%lf\n", N, r);
            } 
        } // end for loop under a specific r
        N = N_int; // set N to initial value to prepare for a new loop with the next r
        r = r + 3.5/i_max; // calculate the next r
    }
    return 0;
}
```
Download [bifurcation.c](https://minhaskamal.github.io/DownGit/#/home?url=https://github.com/liao961120/local_depend/blob/master/1062_class/eco_model/HW/C6/bifurcation.c){target="_blank"}

### R Code

#### Import Data
```{r}
bifurcation <- 
    readr::read_delim("bifurcation.txt", "\t",
                      escape_double = FALSE, 
                      trim_ws = TRUE)
```

#### Plotting
```{r dpi=1000, out.width="70%", out.height="70%"}
library(ggplot2)
library(latex2exp)
ggplot(bifurcation, aes(x=r, y=N))+
    geom_point(size = 0.01, color="#F8766D", alpha = 0.23)+
    scale_x_continuous(breaks = seq(1,5,0.5))+
    scale_y_continuous(breaks = 0:8)+
    labs(title=TeX("N_{t + 1} = N_t e^{r (1-N_t)}"),
         y=TeX("N_t"))
```