### 2018-05-07 1062_class/Multivariate_Analysis/HW3_factor_analysis/HW3.Rmd
up

---
title: "多變量分析 HW3"
author: "劉昱維, 吳冠瑋, 廖永賦, 謝靖惟, 黃奎鈞"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: header.tex
css: /Users/user/local_depend/style_sheets/style.css
always_allow_html: yes
---

```{r setup, include=FALSE}
if (knitr::is_html_output()) {
    knitr::opts_chunk$set(
	echo = FALSE,
	fig.dim = c(3.6, 3.2),
	message = FALSE,
	warning = FALSE,
	comment = "",
	dev = "svg",
	out.width = "70%"
)
} else {
    knitr::opts_chunk$set(echo = FALSE,
                          message=FALSE,
                          warning=FALSE,
                          comment = "",
                          out.width = '70%')
}
options(digits = 4)
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(mat2tex)
theme <- theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
source("multivariate_fc.R")
```

Q1: 9.12
====================================

```{r Covariance_matrices}
S <- 10^(-3)*matrix(c(11.072, 8.019, 8.160,
                      8.019, 6.417, 6.005,
                      8.160, 6.005, 6.773), 
                    nrow = 3, ncol = 3)

Sn <- (23/24)*S # n = 24, Convert to Sn for M.L.E.
```

```{r results='asis', echo=FALSE}
"$\\mathbf{S_n} = \\frac{23}{24} \\mathbf{S} =" %_% xm(Sn ,5) %_% "$"
```

## (a) Specific Variances

```{r}
L <- c(.1022, .0752, .0765)
LL <- matrix(L, ncol = 1) %*% matrix(L, nrow = 1)
Psi <- diag(Sn - matrix(L, ncol = 1) %*% matrix(L, nrow = 1))
```

$\mathbf{S_n} \approx \hat{\mathbf{L}} \hat{\mathbf{L}}^T + \hat{\mathbf{\Psi}}$, where $diag(\mathbf{S_n}) = diag(\hat{\mathbf{L}} \hat{\mathbf{L}}^T) + diag(\hat{\mathbf{\Psi}})$

Hence, $\hat{\mathbf{\Psi}} = diag(\hat{\mathbf{\Psi}}) = diag(\mathbf{S_n} - \hat{\mathbf{L}} \hat{\mathbf{L}}^T)$

```{r results='asis', echo=FALSE}
"$\\hat{\\mathbf{\\Psi}} = " %_% xm(diag(Psi), 6) %_% "$"
```

## (b) Communalities

$\sigma_{ii} = \ell_{i1}^2 + \ell_{i2}^2 + ~...~ + \ell_{im}^2 + \psi_i$

$h_i^2 = \ell_{i1}^2 + \ell_{i2}^2 + ~...~ + \ell_{im}^2 = `r sum(L^2)`$

## (c) Proportion of variance explained by the factor

```{r}
p <-  sum(L^2) / (diag(Sn) %>% sum())
```

$\frac{s_{11} + s_{22} + ~...~ + s_{pp}}{h_i^2} = `r p`$

## (d) Residual Matrix

```{r}
Residual <- Sn - LL - Psi 
```

$\mathbf{S_n} - \mathbf{\hat{L}}\mathbf{\hat{L}}^T - \mathbf{\hat{\Psi}} = 
```{r results='asis'} 
xm(Residual, 6) %_% "$"
```

Q2: 9.32
====================================
```{r}
library(readr)
library(psych)
bulls <- read_table2("T1-10.dat", col_names = FALSE)
colnames(bulls) <- c("Breed", "SalePr", "YrHgt", 
                     "FtFrBody", "PrctFFB", "Frame",
                     "BkFat", "SaleHt", "SaleWt")
bulls <- bulls[,-c(1,9)]
source("fa_functions.R") # for formating loading mt
```

## (a) S PC
```{r}
fa_S_pc <- principal(bulls, 
              nfactors = 2, 
              rotate = 'varimax', 
              covar = TRUE)
fa_S_pc[["loadings"]][,] %>% round(3) %>%
    format_loading_matrix(criterion = 30, markdown = T) %>%
    kable("markdown",align = 'c', escape = F)
```

## (b) S ML

```{r}
fa_S_ml <- fa(bulls, nfactors=2, 
              rotate="varimax", covar = T,
              fm="ml", scores="regression")
fa_S_ml[["loadings"]][,] %>% round(3) %>%
    format_loading_matrix(criterion = 30, markdown = T) %>%
    kable("markdown",align = 'c', escape = F)
```

## (c) R PC

```{r}
fa_R_pc <- principal(bulls, 
              nfactors = 2, 
              rotate = 'varimax', 
              covar = F)
fa_R_pc[["loadings"]][,] %>% round(3) %>%
    format_loading_matrix(criterion = 0.65, markdown = T) %>%
    kable("markdown",align = 'c', escape = F)
```

## (d) R ML
```{r}
fa_R_ml <- fa(bulls, nfactors=2, 
              rotate="varimax", covar = F, 
              fm="ml", scores="regression")
fa_R_ml[["loadings"]][,] %>% round(3) %>%
    format_loading_matrix(criterion = 0.6, markdown = T) %>%
    kable("markdown",align = 'c', escape = F)
```

## (e) Compare **S** & **R**

## (f) scatter plots of factor2 vs factor1 in (a) and (c)

### (a) 
```{r out.width="40%"}
library(ggplot2)
fa_S_pc[["scores"]] %>% as.data.frame() %>%
    ggplot()+
    geom_point(aes(x=RC1, y=RC2), size=1)+
    labs(x="Factor 1", y="Factor 2", title="FA with S",
         caption="Principle component method")
```

### (c) 
```{r out.width="40%"}
library(ggplot2)
fa_R_pc[["scores"]] %>% as.data.frame() %>%
    ggplot()+
        geom_point(aes(x=RC1, y=RC2), size=1)+
        labs(x="Factor 1", y="Factor 2", title="FA with R",
             caption="Principle component method")
```

Q3
====================================

## (a) Appropriate number of factors 
```{r}
library(readr)
drinks <- read_table2("drinks.DAT", col_names = FALSE, skip = 2)
colnames(drinks) <- c("ID", "BRAND",
                      paste("X",1:10, sep=""))
drinks <- drinks[-nrow(drinks),]
```

```{r out.width="40%"}
library(psych)
fa_R_pc2 <- principal(drinks[,3:12], 
              nfactors = 4, 
              rotate = 'varimax', 
              covar = F)
fa_R_pc2[["values"]] %>% 
    cbind(X=1:10) %>% as.data.frame() %>%
    ggplot()+
        geom_point(aes(x=X, y=.), size=2) +
        scale_x_continuous(breaks = 1:10)+
        labs(x="i", y="eigenvalue",
             title="Scree Plot")
```

By the unity criterion we use 4 factors[^fact].

```{r}
fa_R_pc2[["loadings"]][,] %>% round(3) %>%
    format_loading_matrix(criterion = 0.5, markdown = T) %>%
    kable("markdown",align = 'c', escape = F)
```

[^fact]: Although **factor 4** is less than 1, it's eigenvalue is very close to **factor 3**, hence we retain **factor 4**.

## (b) Label the Factors

## (c) Average factor scores of each brand

```{r index}
br1 <- which(drinks$BRAND == "1") #index for brand 1
br2 <- which(drinks$BRAND == "2")
br3 <- which(drinks$BRAND == "3")
br4 <- which(drinks$BRAND == "4")
br5 <- which(drinks$BRAND == "5")
br6 <- which(drinks$BRAND == "6")
```

```{r Pepsi}
fs1 <- fa_R_pc2[["scores"]][br1,] %>% colMeans() %>% t() %>%
    as.data.frame()
fs2 <- fa_R_pc2[["scores"]][br2,] %>% colMeans() %>% t() %>%
    as.data.frame()
fs3 <- fa_R_pc2[["scores"]][br3,] %>% colMeans() %>% t() %>%
    as.data.frame() 
fs4 <- fa_R_pc2[["scores"]][br4,] %>% colMeans() %>% t() %>%
    as.data.frame()
fs5 <- fa_R_pc2[["scores"]][br5,] %>% colMeans() %>% t() %>%
    as.data.frame()
fs6 <- fa_R_pc2[["scores"]][br6,] %>% colMeans() %>% t() %>%
    as.data.frame()

avg_fs <- rbind(fs1, fs2, fs3, fs4, fs5, fs6) %>% as.data.frame()
rownames(avg_fs) <-  c("Pepsi", "Coke", "Gatorade",
                    "Allsport", "Lipton tea", "Nestea")
```

### Average factor scores
```{r factor-score-table}
kable(avg_fs, "markdown", align = 'c',
      digits = 3,
      col.names= c("F1", "F2", "F3", "F4"),
      row.names = T, 
      caption = "Average Factor Scores of each brand")
```

## (d)

```{r}
data <- cbind(fa_R_pc2[["scores"]], as.factor(drinks$BRAND)) %>%
    as.data.frame()
colnames(data) <- c("F1", "F2", "F3", "F4", "Brand")

data[data$Brand =="1", "Brand"] <- "Pepsi"
data[data$Brand =="2", "Brand"] <- "Coke"
data[data$Brand =="3", "Brand"] <- "Gatorade"
data[data$Brand =="4", "Brand"] <- "Allsport"
data[data$Brand =="5", "Brand"] <- "Lipton tea"
data[data$Brand =="6", "Brand"] <- "Nestea"
```

```{r out.width="50%", fig.show='hold'}
### F1, F2
ggplot(data[,c(1,2, 5)]) +
    geom_point(aes(x=F1, y=F2, color=Brand))

### F1, F3
ggplot(data[,c(1,3, 5)]) +
    geom_point(aes(x=F1, y=F3, color=Brand))
```

```{r out.width="50%", fig.show='hold'}
### F1, F4
ggplot(data[,c(1,4, 5)]) +
    geom_point(aes(x=F1, y=F4, color=Brand))

### F2, F3
ggplot(data[,c(2,3, 5)]) +
    geom_point(aes(x=F2, y=F3, color=Brand))
```

```{r out.width="50%", fig.show='hold'}
### F2, F4
ggplot(data[,c(2,4, 5)]) +
    geom_point(aes(x=F2, y=F4, color=Brand))

### F3, F4
ggplot(data[,c(3,4, 5)]) +
    geom_point(aes(x=F3, y=F4, color=Brand))
```