### 2018-05-19 GitHubPage/Instructions/Git command to change local file.Rmd
up

---
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
---

## 1) Clone Repo from GitHub

### Set Working Directory
Set current directory, so your repo can be dowload to this directory: cd \<where to clone>.
```
cd ~/GitHubPage
```

### Clone repo from GitHub
git clone \<URL>
```
git clone https://github.com/liao961120/liao961120.github.io
```

---------
## 2) Working Locally

### Set directory to (cloned) local repo
cd \<REPOSITORY-NAME>
```
cd ~/GitHubPage/liao961120.github.io
```

### (Optional) Create a new branch locally
You can create a branch other than `master` branch locally to work with.

git branch \<BRANCH-NAME>
```
git branch local
```

### Change working branch
After you create a new branch, you could change to that branch (or back), by:

git checkout \<BRANCH-NAME>
```
git checkout master
```

### Sync with Remote
git pull \<remote> \<local branch name>

`git pull` = `git fetch` + `git merge`
```
git pull https://github.com/liao961120/liao961120.github.io master
```

### Edit files locally
You can edit your files in your cloned repo just like you edit other files on your computer.

#### Update changes locally
After editing your files, you need to update changes. Use the git command:

git add \<FILE-NAME>

to update changes you have done to your file.
```
git add ~/GitHubPage/liao961120.github.io/_posts/2017-11-18-markdown-output.md
```

#### Creating new file
If you didn't edit, but want to create a brand new file, create the file with the usual way you do on your computer and use the git command:

git add -f \<FILE-NAME>
```
git add -f ~/GitHubPage/liao961120.github.io/_posts/2017-11-18-markdown-output.md
```

#### Commit change locally
After typing in all the `git add` and/or `git add -f`, type
```
git commit
```
so you can commit locally.

---------

## 3) Update Changes to GitHub

### Upload changes to your GitHub repo
After committing locally, you can update your changes to GitHub.

Use:

git push -u origin \<BRANCH-NAME>

and set \<BRANCH-NAME> to the branch you worked locally.
```
git push -u origin master
```

Now your branch is uploaded to GitHub.

You need to go to GitHub (webpage) to merge your branch to master **only if** you had created and worked on another branch (other than `master`) locally. *See below*.

### (Optional) Merging branch on GitHub

Beware:

* check that you are merging branches under your account

* `base: master` `compare: local`.