### 2018-05-19 GitHubPage/my_R_book/R Learning Notes/01-Basics.Rmd

# Some Basics {#Basic}

## Subsetting Vectors
### vector_name[index vector]
**index vector** can be:

1. logical vectors
2. vectors of positive integers
3. vectors of negative integers
4. vectors of character strings

```{r index_vector1}
#Logical
y <- c(1, -1, 1, 2, 3)
x <- c(NA, 1, 0, -1, 0.5)
sub.y <-  y[y > 0]
sub.x1 <- x[x > 0]
sub.x2 <- x[!is.na(x) & x > 0]
print(list(sub.y=sub.y, sub.x1=sub.x1, sub.x2=sub.x2))
```
*Note: 'NA' is not a value, it's a placeholder for an unknown quantity, the expression NA > 0 evaluates to NA*

```{r index_vector1.2}
#Logical_2: subset with quantile function
x <- runif(100, 0, 1)
df <- as.data.frame(cbind(1:100, x))
names(df) <- c("obs", "value")
df[df$value >= quantile(df$value, 0.95),]
```
-----

```{r index vector2}
# Negative Integer
x <- c("a","b","c","d")
x1 <- x[c(-2, -4)]
x2 <- x[-c(2, 4)]
print(list(x1=x1, x2=x2))
```
*Negative integers give all elements of  x EXCEPT for the 2nd and 10th elements*

-----

```{r index vector3, echo=TRUE}
# Character String
z <- c(foo = 11, bar = 12, norf = NA)
z1 <- z["bar"]
z2 <- z[2]
print(list(z1, z2))
```

## Subsetting Lists
```{r lists, echo=TRUE}
# Return a sublist
ll <- list(id=1:3, height="fdklj", weight=1)
x1 <- ll[1]
x2 <- ll["id"]
print(c(x1, x2))
```
`ll[1]` & `ll["id"]` are equivalent.

------

```{r lists2, echo=TRUE}
# Return a vector(contents in the sublist)
ll <- list(id=1:3, height="fdklj", weight=1)
x1 <- ll[[1]]
x2 <- ll[["id"]]
x3 <- ll$id
print(c(x1,x2,x3))
```
`ll[[1]]`, `ll[["id"]]` *&* `ll$id` *are equivalent.*

*Note:* `[[ ]]` *&* `$` *must go with strings(names).*

Manipulating data with Lists is often difficult. Using `[[ ]]` & `$` lets you manipulate the elements in the List(i.e. vectors).

## Concatinating & Printing
### paste()
The `paste` function combines different vectors(more than 2 vectors as input) into **one** character string. The value of the argument `sep` is used to seperate the inputs.
```{r paste 1, echo=TRUE}
symbol <- "A B C"
num <- 1                           # a numeric vector
x <- paste(symbol, num, sep="\n")  # \n: new line in regex
cat(x)                             # cat is similar to print
```

When the input vector(s) contains multiple elements, and given the `collapse` argument, `paste` collapses the input vector into a single string, seperated by the value of `collapse`. **Beware the element-wise operation(recycle rule)**.
```{r paste 2, echo=TRUE}
x <- 1:3
y <- c("x","y","z")
paste(x,y, sep=":",collapse = ", ")
```
The result of `paste` follows the pattern: `"x[1]:y[1], x[2]:y[2], x[3]:y[3]"`

### cat()