### 2018-05-19 GitHubPage/my_R_book/R Learning Notes/03-Manipulation.Rmd

# Data Manipulation {#Manipulate}

## Explore Data
### View()
View data in new window, `sleep` is a built-in data in R.
```{r, echo=TRUE}
View(sleep)
```
*Note the capital 'V' in `View`.*

### with()
The two lines of code below are equvalent.
```{r with, echo=TRUE}
x1 <- with(airquality,cor(Temp, Wind))
x2 <- cor(airquality$Temp, airquality$Wind)
print(c(x1, x2))
```

### aggragate(y~x, data, function)

Split Temp by Month in data(airquality), then compute mean of Temp split by Month.
```{r aggregate, echo=TRUE}
aggregate(Temp~Month, airquality, mean)
```

### names()
A function to get & change names of an object.
```{r names, echo=TRUE}
x <- c(1, 2, 3)
y1 <- names(x)  # evaluates to NULL
names(x) <- c('a', 'b', 'c')
y2 <- names(x)
print(list(y1=y1, y2=y2))
```

## Handle with missing values
### is<span>.na()
The function returns logical: "TRUE" or "FALSE", indicating whether each element in the data is an 'NA'.
*NA(Not Available) is not a value, but a placeholder for a quantity not available. Hence, **No logical operations with NAs !!!***

R represents 'TRUE' as the number 1 and 'FALSE' as the number 0. Hence, for example, `sum(is.na(sleep))` returns an integer indicating the number of 'NA'.

#### Example: Removing NAs
```{r remove NA1, echo=TRUE}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x)
x.NA_removed <- x[!bad]
print(list(bad=bad, NA_removed=x.NA_removed))
```

### complete.cases()
The function returns TRUE if the row(in data.frame sense) of the data has no NA.
```{r remove NA2}
# Vector example
x <- c(1, NA, 2)
y <- c("a", "b", NA)
good <- complete.cases(x, y)
good
```

#### Example: Remove rows which contains NAs in data frames
```{r airquality head, echo=FALSE}
head(airquality)
```
```{r remove NA3, echo=TRUE}
good <- complete.cases(airquality)
head(airquality[good,])
```
*Note: `good` is a logical vector containing 153 entries(153 rows from airquality)*

## strings

### regular expressions
| Code   | Meaning   |
| :---:  | --------- |
| \\     | excape    |
| \\w    | a-e & 0-9 |
| \\.    | .         |