### 2018-05-19 GitHubPage/my_R_book/R Learning Notes/Chi-Square.Rmd

---
title: "Chi-square"
author: "Yongfu, Liao"
date: "2018年5月3日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## chi-square

### test of goodness-of-fit

Let there be 3 types of flower A, B, C, and we want to know whether the 3 types of flower follow a specific distribution, let it be P(A) = 1/6, P(B) = 1/3, P(C) = 1/2.

To test my hypothesis, 60 flowers were collected, and their types recorded, with 12, 18, 30 of each types (A, B, C), respectively.

```{r}
# chisq.test(x, p)
chisq.test(c(12, 18, 30), p = c(1/6, 1/3, 1/2))
```

### test of independence

```{r}
library(dplyr)
contin_table <- tibble(
    male = c(50, 78),
    female = c(65, 56)) %>% 
    as.data.frame()

rownames(contin_table) <- c("pizza", "cake")
```

```{r}
# chisq.test(contingency_table)
chisq.test(contin_table)
```

#### `table`

```{r}
fct <- c("A", "A", "B", "B","A", "C", "B")
table(fct)
```

```{r}
fq <- table(fct) %>% as.vector()
fq
```

```{r}
fct_name <- table(fct) %>% names()
fct_name
```

```{r}
fq_table <- as.data.frame(fq) %>%
    cbind(fct_name)
fq_table
```