### 2018-05-19 GitHubPage/my_R_book/R Learning Notes/Life table example/Tidy data.Rmd

---
title: "Constructing Life Tables Using R"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have been using the package `dplyr` to handle with data for a while, and I thought I can use it with ease until I was stuck with my homework on **contructing a life table**. I found it easy to do it with spreadsheets, and was embarrassed that I spent that much time dealing it with R. It was due to my unfamiliarity with some functions in R and also insufficient practice.

I will illustrate how I did my homework with R, and you'll find out how easy it is.

## Load Packages

You need these packages to  construct a Life table.
```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(knitr)
```

## Load data
Load the csv file to `life_table`. The raw data contains 3 columns: **Age**, **Survivorship at Age x (lx)**, and **Fecundity at Age x (mx)**. I added `options(scipen=999)` because seeing scientific notations annoys me.
```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA}
options(scipen=999)  # Disable Scientific Notation
life_table <- read_csv("life_table.csv")
life_table
```

## Variables to Construct

Here are the variables that need to be calculated.

|               Statistic               |      Notation     |                         Calculation Formula                        |
|:--------------------------------:|:-----------------:|:------------------------------------------------------------------:|
|                                  |     $l_x m_x$     |                                                                    |
|                                  |    $x l_x m_x$    |                                                                    |
|                                  | $l_x m_x e^{-rx}$ |                                                                    |
| Average survivorship<br>(age class) |       $L_x$       |                      $L_x = (l_x + l_{x+1})/2$                     |
|          Life expectancy         |       $e_x$       |              $e_x = (L_x + L_{x+1} + … + L_{max})/l_x$             |
|        Reproductive value        |       $V_x$       | $\displaystyle \frac{\sum_{y=x}^{max\hspace{0.3mm}x} e^{-ry} l_y m_y}{e^{-rx} l_x}$ |
|       Net reproductive rate      |       $R_0$       |                       $\sum_{all \hspace{0.3mm} x} l_x m_x$                       |
|          Generation time         |        $G$        |                $\frac{\sum_{all \hspace{0.3mm} x} x l_x m_x}{R_0}$                |
|    Intrinsic rate of increase    |  Approximate $r$  |                    $r \approx \frac{ln(R_0)}{G}$                   |
|    Intrinsic rate of increase    |      (True) $r$     |                  $\displaystyle \sum_{all \hspace{0.3mm} x} e^{-rx}l_x m_x = 1$                 |

$l_x*m_x*e^{-rx}$ and $V_x$ will be calculated twice for the approximate and the true $r$.

## Constructing Variables

### `mutate`: Creating new columns

Using the pipe `%>%` and the function `mutate` in package `dplyr`, I first construct 7 new variables. Note the dependencies of the variables, so that I can't construct the life tables just at once.
```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA}
life_table <- life_table %>%
    mutate("lx*mx"=lx*mx,
           "x*lx*mx"=Age*lx*mx,
           "Lx"=(lx+lead(lx))/2,
           "Lx"=replace(Lx, 10, 0),
           "ex"=rev(cumsum(rev(Lx)))/lx,
           "R0"=sum(lx*mx),
           "G"=sum(Age*lx*mx)/R0,
           "approx.r"=log(R0)/G
           )
```

Two things worth noting in the `mutate` function:

1. The code `"Lx"=(lx+lead(lx))/2, "Lx"=replace(Lx, 10, 0)`
    * `lead(lx)` shifts the whole column of lx to its next value, i.e. the column lx becomes lx+1.
    * Due to `lead(lx)`, the last entry of the new column `Lx` must be a `NA`, so I have to assign `0` to it (otherwise all calculations based on it will become `NA`s).

2. The code `"ex"=rev(cumsum(rev(Lx)))/lx` (This is where I was stuck)
    1. The numerator of `ex` is calculated by summing over $L_x$ to $L_{max}$, the maximum age of $L_x$. This is not so intuitive when working with R. 
    2. `rev(Lx)` reverse the order of $L_x$, and `cumsum()` is for *cummulative sum*. `cumsum(rev(Lx))` then is equivalent to summing $L_x$ backwards (i.e. from $L_{max}$ to $L_x$). 
    3. But since $L_x$ is reversed in the first place, `cumsum(rev(Lx))` is also in reverse order. Reversing `cumsum(rev(Lx))` with `rev()` then gives what I want.

### Calculating $r$ by while loop

By the **Eular-Lotka equation**, I can calculate $r$.
$$\displaystyle \sum_{all \hspace{0.3mm} x} e^{-rx}l_x m_x = 1$$

Using `while` loop and Approximate $r$ calculated earlier as the starting value for `r`, $r$ is calculated below. 
```{r, comment=NA}
df <- as.data.frame(life_table)
r <- 0.0812198
x <- sum(exp(-r*df$Age)*df$`lx*mx`)
while (abs(x-1) >= 0.000001) {
    if (x-1>0){
        r <- r+0.00000001
    }
    else{
        r <- r-0.00000001
    }
    x <- sum(exp(-r*df$Age)*df$`lx*mx`)
    r
}
```

The rest is simple, and the logic applied is the same.
```{r, message=FALSE, warning=FALSE, comment=NA}
life_table <- life_table %>%
    mutate(
        "approx.r"=log(R0)/G,
        "r"=r,
        "Vx"=rev(cumsum(rev(exp(-r*Age)*lx*mx)))/exp(-r*Age)*lx,
        "lx*mx*e^-rx"= lx*mx*exp(-r*Age),
        "approx.Vx"=rev(cumsum(rev(exp(-approx.r*Age)*lx*mx)))/exp(-approx.r*Age)*lx,
        "approx.lx*mx*e^-rx"= lx*mx*exp(-approx.r*Age)
    )
```
Now, the life table is done.

## Printing pretty Life Table

```{r, echo=TRUE}
kable(life_table,format="markdown",align="c")
```