### 2018-05-19 GitHubPage/my_R_book/graphics/ggplot.Rmd

---
title: "ggplot quick guide"
author: "Yongfu, Liao"
date: "`r Sys.Date()`"
output: 
    html_document:
        code_folding: show
css: /Users/user/local_depend/style_sheets/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	comment = "",
	dev = 'svg'
)
```

ggplot color theme
--------------------------------
```{r}
scales::hue_pal()(3)
```

Data 
----------------------------------
```{r}
library(ellipse)
library(dplyr)

All_contours <- c(NA, NA, NA) ## Set empty start for appending ##

ell_data <- ellipse(0.8, c(1, 1), centre = c(1, 3), level = 0.5, npoints = 800) ## npoints: bigger contours with more points ##

ell_data <- as.data.frame(ell_data) %>%
    mutate(class = "50% contour")

## Change to data.frame BEFORE cbind, or coersion happens ##

All_contours <- rbind(All_contours, ell_data)[-1,]
```

Adding Lines & Curves to a plot
-------------------------------------

```{r}
library(latex2exp)
library(ggplot2)

sum_xy <- All_contours %>%
    mutate(sum=x+y) %>% select(sum)

max <- which(sum_xy == max(sum_xy))[1]
min <- which(sum_xy == min(sum_xy))[1]

p <- ggplot(data = All_contours) +
    geom_point(aes(x = x, y = y, color = class),
               size = 0.1,
               color="black")+
    labs(color = "Contours", 
         title = "Contour Plot",
         x = TeX("$x_1$"), y = TeX("$x_2$")
    )

p + geom_segment(aes(x=All_contours$x[max], 
                     y=All_contours$y[max], 
                     xend =All_contours$x[min], 
                     yend = All_contours$y[min], 
                     color = "Major Axis")) +
    geom_segment(aes(x=0.6342645, 
                     y=3.378911, 
                     xend =1.370133, 
                     yend = 2.625475, 
                     color = "Minor Axis")) +
    annotate("text", x = 1.5, y = 3.7, 
             label =TeX("$c\\sqrt{\\lambda_1}$"),
             alpha =0.7,
             size = 2.8,
             color = "red")+
    coord_fixed() # fix x-axis/y-axis to 1
```