### 2018-06-14 1062_class/Multivariate_Analysis/HW5_Clustering/multivarite_hw4.Rmd
up

---
title: "多變量分析 HW4"
author: "劉昱維, 吳冠瑋, 廖永賦, 謝靖惟"
css: /home/liao/local_depend/style_sheets/style.css
output:
  bookdown::pdf_document2:
    number_sections: no
    includes:
      in_header: /home/liao/local_depend/pdf_document/header.tex
    latex_engine: xelatex
  bookdown::html_document2:
    includes:
      before_body: mathjax_bm.tex
    number_sections: no
geometry: margin=0.7in
fontsize: 11pt
always_allow_html: yes
---

```{r setup, include=FALSE}
library(svglite)
if (knitr::is_html_output()) {
    knitr::opts_chunk$set(
	echo = FALSE,
	fig.dim = c(7, 4),
	message = FALSE,
	warning = FALSE,
	comment = "",
	dev='svg',
	out.width = "100%"
)
} else {
    library(showtext)
    knitr::opts_chunk$set(echo = FALSE,
                          message=FALSE,
                          warning=FALSE,
                          comment = "",
                          out.width = '70%',
                          fig.showtext = T)
}
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)
library(mat2tex)
theme <- theme_light() +
    theme(axis.text.x = element_text(size = 7, face = "plain", angle = 30),
               axis.text.y = element_text(size = 7, face = "plain"),
               axis.title.x = element_text(size = 9, face = "bold"),
    axis.title.y = element_text(size = 9, face = "bold"))
```