### 2018-07-10 GitHubPage/my_R_book/Rmarkdown_features/beamer/beamer.Rmd
up

---
title: "Create Beamer Slide with R Markdown"
subtitle: "A Reproducible Example"
author: "Yongfu, Liao"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
institute: "National Taiwan University"
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
    fonttheme: "structurebold"
    incremental: false
    latex_engine: xelatex
    includes:
        in_header: header_beamer.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
date_text <- "`r format(Sys.Date(), '%B %d, %Y')`"
```

## Reproduce This slide: yaml Header

```yaml
---
title: "Create Beamer Slide with R Markdown"
subtitle: "A Reproducible Example"
author: "Yongfu, Liao"
date: "`r date_text`"
institute: "National Taiwan University"
```

## Reproduce This slide: yaml Header (Cont.)
```yaml
fontsize: 9pt
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
    fonttheme: "structurebold"
    incremental: false
    latex_engine: xelatex
    includes:
        in_header: header_beamer.tex
---
```

- Links
    - [header_beamer.tex](https://github.com/liao961120/local_depend/blob/master/style_sheets/header_beamer.tex)
    - [Source Directory](https://github.com/liao961120/local_depend/tree/master/GitHubPage/my_R_book/Rmarkdown_features/beamer) of this Presentation
    - [More](https://liao961120.github.io/notes/rmd_features.html) about R Markdown

## Two-Column Layout

\begin{columns}
\column{0.75\textwidth}

```{r out.width="75%"}
library(ggplot2)
ggplot(iris, aes(x = Sepal.Width, y = Sepal.Length)) +
    geom_point(aes(color=Species), show.legend = F)
```

\column{0.25\textwidth}
R is a collaborative project with many contributors.
Type `contributors()` for more information and
`citation()` on how to cite R or R packages in publications.
\end{columns}

## Two-Column Layout: Code

```tex
\begin{columns}
\column{0.75\textwidth}
```
    `r ''````{r out.width="75%"}
    library(ggplot2)
    ggplot(iris, aes(x = Sepal.Width,
                     y = Sepal.Length)) +
        geom_point(aes(color = Species),
                   show.legend = F)
    ```
```tex
\column{0.25\textwidth}
R is a collaborative project with many contributors.
Type `contributors()` for more information and
`citation()` on how to cite R or R packages in publications.
\end{columns}
```

## Various Blocks

\begin{alertblock}{Red Block}
$\frac{\int_{-\infty}^{\infty} f(x)dx}{b}$
\end{alertblock}

\begin{examples}
Example Blocks are green.
\end{examples}

### Blue Block
Paragraph 1

Paragraph 2

## Various Blocks: Code

**Red Block**: Can set titles
```tex
\begin{alertblock}{Red Block}
$\frac{\int_{-\infty}^{\infty} f(x)dx}{b}$
\end{alertblock}
```

**Examples**: Can't change title
```tex
\begin{examples}
Example Blocks are green.
\end{examples}
```

**Blue Block**: h3 level Headings[^h3]
```markdown
### Blue Block
Paragraph 1: Text In Box

Paragraph 2: Blahblahblah
```

[^h3]: Note this only works in 'Madrid' theme in R markdown.

## Incremental Bullets

> 1. Item 1

> 2. Item 2

> 3. Item 3

## Contact

```{r out.width="10%"}
knitr::include_graphics("tux.png")
```

\

Where You can find me:

- [GitHub](https://github.com/liao961120)

- [Personal Page](https://liao961120.github.io/)