### 2018-07-14 Ubuntu_16.04_setup/mylinux.Rmd
up

---
title: "My Linux Notes"
subtitle: "A Quick Reference"
author: "[Yongfu Liao](https://liao961120.github.io/)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
link-citations: yes
output:
  bookdown::html_document2:
    css: /home/liao/local_depend/style_sheets/style.css
    df_print: kable
    toc: true
    toc_depth: 2
    toc_float:
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg',
	fig.dim = c(5.5, 3.5)  # (width, height)
)
```

Navigation
=====

## Soft Link

Syntax: `ln -s <target> <link-name>`

e.g.
```bash
ln -s $(pwd)/blah.txt ~/Desktop/blah-sym
```

Network
=====

## Setup SSH

### Server Setup{-}
Install SSH Server (so be connected by a remote client):
```bash
sudo apt-get install openssh-server
```

Configure:
```bash
sudo gedit /etc/ssh/sshd_config
```

> Port 22
> PasswordAuthentication yes
> PermitRootLogin yes

Findout IP address locally: `hostname -I`

### Client Setup{-}

Access from remote via SSH:

Syntax: `ssh <server-user-account>@<ip>`

e.g.
```bash
ssh liao961120@192.168.2.110
```

Remote Control Via SSH
=====

## Hardware Configuration

1. Fix floating IP
(Alternative: auto push IP address when turned on
https://superuser.com/questions/471253/ubuntu-how-to-run-a-terminal-command-on-boot)
2. Enable WakeOnLane(WOL) at BIOS: 
	1. `Power Management Setup` > `WOL: enabled`[^wol]
	2. `Power Management Setup` > `deep power off mode: disabled`[^poff]

[^wol]: There may be different names for WOL, such as 'Wake Up on Lan', 'Automatic Power Up', 'Resume by LAN', 'Power Up On PCI Device', 'PME Event Wake Up', etc.
[^poff]: This keeps the power for NIS to turn on the computer when receiving a signal from web.
[^utc]: Note the date time is in [Coordinated Universal Time (UTC) Standard](https://en.wikipedia.org/wiki/Coordinated_Universal_Time), so it's -8 hrs from the time in Taiwan.

## wakeonlan

wakeonlan -i <broadcast-ip> <MAC>

```bash
wakeonlan -i 192.168.2.255 44:87:fc:a1:0b:cc
wakeonlan -i $rm_ip_b $rm_MAC
```

## SSH login

ssh liao961120@<ip>

```bash
ssh liao961120@192.168.2.110
ssh $ssh_pc
```

## SSH logout & suspend computer

Use `sudo shutdown 0` to leave. Remember to enable `Power of Mode` in BIOS, so you can WOL to open the computer again.

## Variable Definition (in `~/.bashrc`)
```bash
rm_MAC="44:87:fc:a1:0b:cc"
rm_ip="192.168.2.110"
rm_ip_b=`bash ~/ip2ipb.sh $rm_ip`
ssh_pc="liao961120@$rm_ip"
```