### 2018-10-15 GitHubPage/my_R_book/Rmarkdown_features/03-slide.rmd

---
title: "R Markdown Features"
subtitle: "A Compact Reference"
author: "[Yongfu Liao](https://liao961120.github.io/)"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
bibliography: /home/liao/local_depend/citation/neanderthals.bib
csl: /home/liao/local_depend/style_sheets/apa.csl
link-citations: yes
output:
  bookdown::html_document2:
    css: style.css
    df_print: kable
    toc: true
    toc_depth: 2
    toc_float:
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg',
	fig.dim = c(5.5, 3.5)  # (width, height)
)
str_date <- "`r format(Sys.Date(), '%B %e, %Y')`"
```

# Xaringan (remark.js Slide)

Xaringan is a HTML presentation output format of R Markdown. Xaringan doesn't support Pandoc markdown, but is very flexible and easy to use if having basic CSS knowledge. It also prints well on PDF (using Chrome or [decktape](https://github.com/astefanutti/decktape)). Read [Chapter 7](https://bookdown.org/yihui/rmarkdown/xaringan.html) of *R Markdown Guide* for more details.

<iframe src="https://liao961120.github.io/slides/xaringan/index.html", width=400, height=301>
</iframe>  
[Slide Link](https://liao961120.github.io/slides/xaringan/index.html)

# Beamer Slide

Beamer slide is popular among proficient LaTeX users. However, I think Xaringan is a better option since it's able to print to PDF properly and at the same time having the ability to use interactive HTML features[^xaringan].

<embed src="beamer.pdf" type="application/pdf" width="90%" height="420px">

Note: You can use (left & right) arrow keys to navigate through the slides page-by-page.

[^xaringan]: An automatic process can be set to render Xaringan slides, print them to PDF, and host HTML and PDF slides on GitHub Pages. See the [output webpage](/slides) and this [.travis.yml](https://github.com/liao961120/slides/blob/master/.travis.yml) if you want to implement it.