### 2018-10-23 GitHubPage/my_R_book/Rmarkdown_features/00-yaml.rmd
up

---
title: "R Markdown Features"
subtitle: "A Compact Reference"
author: "[Yongfu Liao](https://liao961120.github.io/)"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
bibliography: /home/liao/local_depend/citation/neanderthals.bib
csl: /home/liao/local_depend/style_sheets/apa.csl
link-citations: yes
output:
  bookdown::html_document2:
    css: style.css
    df_print: kable
    toc: true
    toc_depth: 2
    toc_float:
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg',
	fig.dim = c(5.5, 3.5)  # (width, height)
)
str_date <- "`r format(Sys.Date(), '%B %e, %Y')`"
```

# yaml Setup {.tabset}

yaml metadata determines the output file format, looking, and any other additional customizing features of Rmarkdown files.

## yaml {-}

This is the complete yaml header of this document.

```{yaml}
---
title: "R Markdown Features"
subtitle: "A Compact Reference"
author: "[Yongfu Liao](https://liao961120.github.io/)"
date: "`r format(Sys.Date(), '%B %e, %Y')`" 
bibliography: /home/liao/local_depend/citation/neanderthals.bib
csl: /home/liao/local_depend/style_sheets/apa.csl
link-citations: yes
output:
  bookdown::html_document2:
    code_folding: show
    css: /home/liao/local_depend/style_sheets/style.css
    df_print: kable
    toc: true
    toc_depth: 2
    toc_float:
        collapsed: false
---
```

## HTML {-}

```{yaml}
output:
  html_document:
    includes:
        before_body: mathjax_bm.tex
    css: ../style_sheets/style.css
    df_print: kable
    code_folding: hide
    fig_width: 7
    fig_height: 6
    toc: true
    toc_depth: 4
    toc_float: true
```

## PDF {-}

You may need to attatch an additional `header.tex` and set `latex_engine` to `xelatex` for PDF documents containing multibite characters (e.g. Chinese). Read more about  [header.txt](#pdf-header).
```{yaml}
---
author: "廖永賦\\thanks{組別：3 ~~ 系級：心理三 ~~ 學號：B00000000}"
output:
  pdf_document:
    latex_engine: xelatex    
    includes:
        in_header: header.tex
documentclass: article
urlcolor: "blue"
linestretch: 1.5
fontsize: 12pt 
geometry: a5paper, headheight=0pt, margin=0.4in
linestretch: 1.5
links-as-notes: true
---
```

Note that Latex only accepts fontsize of 10pt, 11pt, or 12pt. See [this](https://github.com/rstudio/rmarkdown/issues/878) for more detail.

## PDF 2 {-}
This is the yaml header for `bookdown::pdf_document2` output. Using `bookdown` extensions have several advantages. Search 'bookdown' for details.
```{yaml}
---
author: "廖永賦\\thanks{組別：3 ~~ 系級：心理三 ~~ 學號：B00000000}"
output:
  bookdown::pdf_document2:
    number_sections: FALSE
    toc: False
    includes:
      in_header: header.tex
    latex_engine: xelatex
  bookdown::html_document2:
    number_sections: FALSE
    includes:
      before_body: mathjax_bm.tex
urlcolor: blue
always_allow_html: yes
linestretch: 1.5
geometry: a4paper #, headheight=0pt, margin=0.4in, landscape
links-as-notes: true
---
```

## PDF: header.tex{- #pdf-header}

The code below is for generating PDF documents written in (Traditional) Chinese.

```tex
`r paste(readLines("~/local_depend/style_sheets/header.tex"), collapse = '\n')`
```

## PDF: template {- #pdf-template}

```markdown
output:
  pdf_document:
    template: two-col.latex
    includes:
      in_header: header.tex
```

It is possible to define [custom latex template](https://rmarkdown.rstudio.com/pdf_document_format#custom_templates).
For example, [`two-col.latex`](https://github.com/liao961120/local_depend/blob/master/style_sheets/two-col.latex) is created by putting the variable `body` (which is the rendered content of the Rmd file) inside `multicols` latex extension:

```tex
\begin{multicols}{2}
$body$
\end{multicols}
```

## Bibliography {-}

```{yaml}
bibliography: ./citation/neanderthals.bib
csl: ./citation/apa.csl # APA style sheet
link-citations: yes
```

For citation syntax, read the section [Bibliography & Citations](#cite).