### 2018-12-11 GitHubPage/my_R_book/Rmarkdown_features/03-slide.rmd

# (PART) **Presentation** {-}

# Xaringan (remark.js Slide)

Xaringan is a HTML presentation output format of R Markdown. Xaringan doesn't support Pandoc markdown, but is very flexible and easy to use if having basic CSS knowledge. It also prints well on PDF (using Chrome or [decktape](https://github.com/astefanutti/decktape)). Read [Chapter 7](https://bookdown.org/yihui/rmarkdown/xaringan.html) of *R Markdown Guide* for more details.

<iframe src="https://liao961120.github.io/slides/xaringan/index.html", width=400, height=301>
</iframe>  
[Slide Link](https://liao961120.github.io/slides/xaringan/index.html)

# Beamer Slide

Beamer slide is popular among proficient LaTeX users. However, I think Xaringan is a better option since it's able to print to PDF properly and at the same time having the ability to use interactive HTML features[^xaringan].

<embed src="beamer.pdf" type="application/pdf" width="90%" height="420px">

Note: You can use (left & right) arrow keys to navigate through the slides page-by-page.

[^xaringan]: An automatic process can be set to render Xaringan slides, print them to PDF, and host HTML and PDF slides on GitHub Pages. See the [output webpage](/slides) and this [.travis.yml](https://github.com/liao961120/slides/blob/master/.travis.yml) if you want to implement it.