### 2018-12-11 GitHubPage/my_R_book/Rmarkdown_features/index.rmd

---
title: "R Markdown Features"
subtitle: "A Compact Reference"
author: "[Yongfu Liao](https://liao961120.github.io/)"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
bibliography: /home/liao/local_depend/citation/neanderthals.bib
csl: /home/liao/local_depend/style_sheets/apa.csl
link-citations: yes
output:
  bookdown::gitbook:
    split_by: rmd
    df_print: kable
    toc_depth: 1
    #self_contained: true
    output_dir: _book
---

```{r setup, include=FALSE}
options(knitr.duplicate.label = 'allow')
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev='svg',
	fig.dim = c(5.5, 3.5)   # width height
	)
str_date <- "`r format(Sys.Date(), '%B %e, %Y')`"
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'), all_precode = T)
```